---
title: "Stat 477 - HW 3"
author: "Jay Maxwell"
date: "2/8/2022"
output: 
  pdf_document:
    fig_caption: yes
    includes:
     in_header: my_header.tex
---  

```{r setup, include=FALSE, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(tidy.opts=list(width.cutoff=80), 
                      tidy=TRUE, 
                      include=TRUE, 
                      echo=TRUE) 
                      #results="asis",
                      #fig.pos = "!", 
                      #out.extra = ""
library(plyr)                      
library(tidyverse)
library(kableExtra)
library(knitr)
library(epiDisplay)

## FOLLOWING CODE PROVIDED BY P. CARAGEA, IOWA STATE UNIV.
##Enter the function prop.ci used to calculate the 
##confidence interval for a population proportion using either 
##the Wilson score method or normal approximation method
  prop.ci<- function(y, n, type, conf.level){
    phat<- y/n
    alpha<- 1 - conf.level
    z<- qnorm(1 - alpha/2)
    newy<- y + z^2/2
    newn<- n + z^2
    newphat<- newy/newn
  
    if (type=="normal") {
      lowerci<- phat - z*sqrt(phat*(1-phat)/n)
      upperci<- phat + z*sqrt(phat*(1-phat)/n)
    }
  
    else if (type=="score") {
      lowerci<- (phat + (1/(2*n))*z^2 - z*sqrt(phat*(1-phat)/n + z^2/(4*n^2)))/(1 + (1/n)*z^2)
      upperci<- (phat + (1/(2*n))*z^2 + z*sqrt(phat*(1-phat)/n + z^2/(4*n^2)))/(1 + (1/n)*z^2)
    }
  
    else {
      lowerci<- newphat - z*sqrt(newphat*(1-newphat)/newn)
      upperci<- newphat + z*sqrt(newphat*(1-newphat)/newn)
    }
    cat(lowerci, upperci, "\n")
  }

##Enter the function nprop.ci used to calculate the sample size
##necessary for a confidence interval with a given margin of 
##error and confidence level
  nprop.ci<- function(p, m, conf.level) {
    alpha<- 1 - conf.level
    z<- qnorm(1 - alpha/2)
    ceiling((z/m)^2*p*(1-p))
  }

#############################################################################################################

  coverage.ci<- function(n, p, type, conf.level){
  sample<- rbinom(100000, n, p)
  
  phat<- sample/n
  alpha<- 1 - conf.level
  z<- qnorm(1 - alpha/2)
  newy<- sample + z^2/2
  newn<- n + z^2
  newphat<- newy/newn
  
  lowerci<- rep(0, 100000)
  upperci<- rep(0, 100000)
  nocov<- rep(0, 100000)
  
  for (i in 1:100000){
    if (type=="normal") {
      lowerci[i]<- phat[i] - z*sqrt(phat[i]*(1-phat[i])/n)
      upperci[i]<- phat[i] + z*sqrt(phat[i]*(1-phat[i])/n)
    }
    else if (type=="score") {
      lowerci[i]<- (phat[i] + (1/(2*n))*z^2 - z*sqrt(phat[i]*(1-phat[i])/n + z^2/(4*n^2)))/(1 + (1/n)*z^2)
      upperci[i]<- (phat[i] + (1/(2*n))*z^2 + z*sqrt(phat[i]*(1-phat[i])/n + z^2/(4*n^2)))/(1 + (1/n)*z^2)
    }
    else {
      lowerci[i]<- newphat[i] - z*sqrt(newphat[i]*(1-newphat[i])/newn)
      upperci[i]<- newphat[i] + z*sqrt(newphat[i]*(1-newphat[i])/newn)
    }
    nocov[i]<- ifelse(lowerci[i] > p, 1, 0) + ifelse(upperci[i] < p, 1, 0)
  }
  covrate<- 1 - sum(nocov)/100000
  #cat(covrate*100, "\n")
  covrate*100
}
  
## END P CARAGEA CODE
```

1. According to a Gallup poll, of 1,000 randomly selected adults aged 18 or older in the United States, 65% believe that global warming is more a result of human actions than natural causes.

|      (a) Use R to obtain a summary table and bar graph of the sample data.

```{r q1a}

warming <- as.factor(c(rep('yes', 650), rep('no',350)))
df <- data.frame(warming)
tab1(df$warming, graph=FALSE, cum.percent = FALSE)

df %>% ggplot(aes(x=warming))+
    geom_bar(fill = "steelblue", colour = "black") +
    ylim(0, 800) +
    labs(x = "Belief",
         y = "Number of People",
         title = "Belief that global Warming is caused by people") +
    theme_bw() +
    theme(axis.title.y = element_text(size = rel(1.2)),
          axis.title.x = element_text(size = rel(1.2)),
          axis.text.x = element_text(size = rel(1.4)),
          axis.text.y = element_text(size = rel(1.4)),
          plot.title = element_text(hjust=0.5, size = rel(1.5)))
```



|      (b) Describe the population proportion of interest p in words.

**In this example, the population proportion (p) is the faction of the entire US population that believes global warming is caused more by people than natural causes. **



|      (c) Use R to calculate a 95% confidence interval for the population proportion p using the normal approximation method.

```{r}
prop.ci(650,1000,type="normal", 0.95)
```




|      (d) Give the interpretation of the 95% confidence interval you calculated in part (c) in the context of the problem. 

**We are 95% confident that the true population proportion of people in the USE who believe global warming is caused more by people than nature is between 62.04% to approximately 67.95%.**





|      (e) Use R to calculate a 95% confidence interval for the population proportion p using the Wilson’s score method. Compare the center and width of this interval to the one you calculated in part (c).


```{r}
prop.ci(650,1000,type="score", 0.95)


(.6795623+.6204377)/2
(.6789373+.6199147)/2

```

**The results of the score method have a slightly smaller range than the results of the normal approxmation method. I would call the centers for each calculation approximately equal at .65. **



|      (f) Gallup is planning to conduct another poll on global warming. They would like to have a 95% confidence interval with a margin of error of no more than 2.5%. What sample size do they need to obtain this margin of error? Make sure to specify how you are calculating the sample size 

```{r}

cat("Sample size needed:",nprop.ci(.65,0.025,0.95))

```

**To calculate the sample size, we solve for $n\geq (\frac{z_{1-\frac{\alpha}{2}}}{M})^2 \hat{\hat{p}}*(1-\hat{\hat{p}})$. Because this is a follow up survey, the reasearches might use their prior findings to estimate $\hat{\hat{p}}=.65$. **

2. Unlike confidence intervals for other parameters... What do you notice about the coverage rates of the two methods? How do these results
depend on the values of n and p?


\begin{table}[h!]
\begin{tabular}{|l||lll||lll|}
\hline
                            &      & Normal &   & & Wilson's &  \\ \hline
\multicolumn{1}{|c||}{$p$} & n=25 & n=250 & n=1000 & n=25 & n=250 & n=1000 \\ \hline
0.5      & 95.651 & 95.145 & 94.656  & 95.766 & 95.008 & 94.729 \\
0.75     & 89.468 & 93.999 & 94.66 & 93.901 & 94.365 & 94.758 \\
0.9      & 91.851 & 93.104 & 95.17 & 96.33 & 95.54 & 94.893 \\ \hline
\end{tabular}
\end{table}

**When we compare the coverages calcuated above it appears that for most of our simulations the Wilson's score method results in coverage closer to the 95% goal more sore than the Normal approximation method. I have a hard time determining a pattern of relation between n and p for both methods. There is a lot of variation in coverage (betwen 89% and 95%) for the normal method for all the combinations of n and p. While the Wilson's score method seems to stay consistently close to our ideal coverage of 95% no matter what the combinations of n and p are changed to. **

```{r q2, echo=TRUE}
# coverage.ci(25,.5,"normal",0.95)
# coverage.ci(250,.5,"normal",0.95)
# coverage.ci(1000,.5,"normal",0.95)
# 
# coverage.ci(25,.5,"score",0.95)
# coverage.ci(250,.5,"score",0.95)
# coverage.ci(1000,.5,"score",0.95)
# 
# 
# coverage.ci(25,.75,"normal",0.95)
# coverage.ci(250,.75,"normal",0.95)
# coverage.ci(1000,.75,"normal",0.95)
# 
# coverage.ci(25,.75,"score",0.95)
# coverage.ci(250,.75,"score",0.95)
# coverage.ci(1000,.75,"score",0.95)
# 
# coverage.ci(25,.9,"normal",0.95)
# coverage.ci(250,.9,"normal",0.95)
# coverage.ci(1000,.9,"normal",0.95)
# 
# coverage.ci(25,.9,"score",0.95)
# coverage.ci(250,.9,"score",0.95)
# coverage.ci(1000,.9,"score",0.95)
```


